---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your group name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one student name, class group (just the letter; A, B, or C), and ID per line, e.g., 123456 - A - John Leposo; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  ID - Class Group - Name                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 2.  ID - Class Group - Name                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 3.  ID - Class Group - Name                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 4.  ID - Class Group - Name                                                                                                                                              |
|                                                   |                                                                                                                                                                          |
|                                                   | 5.  ID - Class Group - Name                                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | *\<specify the name of the team you created on GitHub classroom\>*                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----
if (!is.element("formatR", installed.packages()[, 1])) {
  install.packages("formatR", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("formatR")


## readr - Load datasets from CSV files ----
if (!is.element("readr", installed.packages()[, 1])) {
  install.packages("readr", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("readr")

if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr ----
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071 ----
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## factoextra ----
if (require("factoextra")) {
  require("factoextra")
} else {
  install.packages("factoextra", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## FactoMineR ----
if (require("FactoMineR")) {
  require("FactoMineR")
} else {
  install.packages("FactoMineR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

# Loading the Student Performance Dataset

The 20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset is then loaded. The dataset and its metadata are available here: <https://drive.google.com/drive/folders/1-BGEhfOwquXF6KKXwcvrx7WuZXuqmW9q?usp=sharing>

```{r Load Dataset}
student_performance_dataset <-
  readr::read_csv(
                  "/Users/Cris/github-classroom/BBT4206-R-Lab4of15-DataTransforms-team-kubwa/data/20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset - dataset.csv", 
                  col_types =
                  readr::cols(
                              class_group =
                              readr::col_factor(levels = c("A", "B", "C")),
                              gender = readr::col_factor(levels = c("1", "0")),
                              YOB = readr::col_date(format = "%Y"),
                              regret_choosing_bi =
                              readr::col_factor(levels = c("1", "0")),
                              drop_bi_now =
                              readr::col_factor(levels = c("1", "0")),
                              motivator =
                              readr::col_factor(levels = c("1", "0")),
                              read_content_before_lecture =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              anticipate_test_questions =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              answer_rhetorical_questions =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              find_terms_I_do_not_know =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              copy_new_terms_in_reading_notebook =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              take_quizzes_and_use_results =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              reorganise_course_outline =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              write_down_important_points =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              space_out_revision =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              studying_in_study_group =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              schedule_appointments =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              goal_oriented =
                              readr::col_factor(levels =
                                                c("1", "0")),
                              spaced_repetition =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              testing_and_active_recall =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              interleaving =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              categorizing =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              retrospective_timetable =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              cornell_notes =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              sq3r = readr::col_factor(levels =
                                                       c("1", "2", "3", "4")),
                              commute = readr::col_factor(levels =
                                                          c("1", "2",
                                                            "3", "4")),
                              study_time = readr::col_factor(levels =
                                                             c("1", "2",
                                                               "3", "4")),
                              repeats_since_Y1 = readr::col_integer(),
                              paid_tuition = readr::col_factor(levels =
                                                               c("0", "1")),
                              free_tuition = readr::col_factor(levels =
                                                               c("0", "1")),
                              extra_curricular = readr::col_factor(levels =
                                                                   c("0", "1")),
                              sports_extra_curricular =
                              readr::col_factor(levels = c("0", "1")),
                              exercise_per_week = readr::col_factor(levels =
                                                                    c("0", "1",
                                                                      "2",
                                                                      "3")),
                              meditate = readr::col_factor(levels =
                                                           c("0", "1",
                                                             "2", "3")),
                              pray = readr::col_factor(levels =
                                                       c("0", "1",
                                                         "2", "3")),
                              internet = readr::col_factor(levels =
                                                           c("0", "1")),
                              laptop = readr::col_factor(levels = c("0", "1")),
                              family_relationships =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              friendships = readr::col_factor(levels =
                                                              c("1", "2", "3",
                                                                "4", "5")),
                              romantic_relationships =
                              readr::col_factor(levels =
                                                c("0", "1", "2", "3", "4")),
                              spiritual_wellnes =
                              readr::col_factor(levels = c("1", "2", "3",
                                                           "4", "5")),
                              financial_wellness =
                              readr::col_factor(levels = c("1", "2", "3",
                                                           "4", "5")),
                              health = readr::col_factor(levels = c("1", "2",
                                                                    "3", "4",
                                                                    "5")),
                              day_out = readr::col_factor(levels = c("0", "1",
                                                                     "2", "3")),
                              night_out = readr::col_factor(levels = c("0",
                                                                       "1", "2",
                                                                       "3")),
                              alcohol_or_narcotics =
                              readr::col_factor(levels = c("0", "1", "2", "3")),
                              mentor = readr::col_factor(levels = c("0", "1")),
                              mentor_meetings = readr::col_factor(levels =
                                                                  c("0", "1",
                                                                    "2", "3")),
                              `Attendance Waiver Granted: 1 = Yes, 0 = No` =
                              readr::col_factor(levels = c("0", "1")),
                              GRADE = readr::col_factor(levels =
                                                        c("A", "B", "C", "D",
                                                          "E"))),
                  locale = readr::locale())
```

## Description of the Dataset

We then display the number of observations and number of variables. We have 101 observations and 100 variables to work with.

```{r Your Fourth Code Chunk}
dim(student_performance_dataset)
```

Next, we display the quartiles for each numeric variable[*... think of this process as **"storytelling using the data."** Tell us what is happening; tell us what you are discovering as you proceed with the markdown; walk us through your code step-by-step (a code walkthrough).*]{#highlight style="color: blue"}

```{r Your Fifth Code Chunk}
summary(student_performance_dataset)
```

# \<Scale Data Transform\>

The scale data transform is useful for scaling data that has a Gaussian distribution. The scale data transform works by calculating the standard deviation of an attribute and then divides each value by the standard deviation.

```{r Your Sixth Code Chunk}
# Fill this with R related code that will be executed when the R markdown file
# BEFORE
class(student_performance_dataset)
summary(student_performance_dataset)
colnames(student_performance_dataset)
student_performance_dataset<- as.data.frame(student_performance_dataset)

hist(student_performance_dataset[, 75], main = names(student_performance_dataset)[75])
hist(student_performance_dataset[, 76], main = names(student_performance_dataset)[76])
hist(student_performance_dataset[, 77], main = names(student_performance_dataset)[77])
hist(student_performance_dataset[, 78], main = names(student_performance_dataset)[78])
hist(student_performance_dataset[, 79], main = names(student_performance_dataset)[79])
hist(student_performance_dataset[, 80], main = names(student_performance_dataset)[80])
hist(student_performance_dataset[, 81], main = names(student_performance_dataset)[81])
hist(student_performance_dataset[, 82], main = names(student_performance_dataset)[82])
hist(student_performance_dataset[, 83], main = names(student_performance_dataset)[83])
hist(student_performance_dataset[, 84], main = names(student_performance_dataset)[84])
hist(student_performance_dataset[, 85], main = names(student_performance_dataset)[85])
hist(student_performance_dataset[, 86], main = names(student_performance_dataset)[86])
hist(student_performance_dataset[, 87], main = names(student_performance_dataset)[87])

model_of_the_transform <- preProcess(student_performance_dataset, method = c("scale"))
print(model_of_the_transform)
student_performance_dataset_scale_transform <- predict(model_of_the_transform,
                                                       student_performance_dataset)

# AFTER comment
summary(student_performance_dataset_scale_transform)
hist(student_performance_dataset_scale_transform[, 75],
     main = names(student_performance_dataset_scale_transform)[75])
hist(student_performance_dataset_scale_transform[, 76],
     main = names(student_performance_dataset_scale_transform)[76])
hist(student_performance_dataset_scale_transform[, 77],
     main = names(student_performance_dataset_scale_transform)[77])
hist(student_performance_dataset_scale_transform[, 78],
     main = names(student_performance_dataset_scale_transform)[78])
hist(student_performance_dataset_scale_transform[, 79],
     main = names(student_performance_dataset_scale_transform)[79])
hist(student_performance_dataset_scale_transform[, 80],
     main = names(student_performance_dataset_scale_transform)[80])
hist(student_performance_dataset_scale_transform[, 81],
     main = names(student_performance_dataset_scale_transform)[81])
hist(student_performance_dataset_scale_transform[, 82],
     main = names(student_performance_dataset_scale_transform)[82])
hist(student_performance_dataset_scale_transform[, 83],
     main = names(student_performance_dataset_scale_transform)[83])
hist(student_performance_dataset_scale_transform[, 84],
     main = names(student_performance_dataset_scale_transform)[84])
hist(student_performance_dataset_scale_transform[, 85],
     main = names(student_performance_dataset_scale_transform)[85])
hist(student_performance_dataset_scale_transform[, 86],
     main = names(student_performance_dataset_scale_transform)[86])
hist(student_performance_dataset_scale_transform[, 87],
     main = names(student_performance_dataset_scale_transform)[87])

# BEFORE
summary(student_performance_dataset)
# The code below converts column number 99 into unlisted and numeric data first
# so that a histogram can be plotted. Further reading:
student_performance_dataset_yield <- as.numeric(unlist(student_performance_dataset[, 99]))
hist(student_performance_dataset_yield, main = names(student_performance_dataset_yield)[99])

model_of_the_transform <- preProcess(student_performance_dataset, method = c("scale"))
print(model_of_the_transform)
student_performance_data_scale_transform <- predict(model_of_the_transform, student_performance_dataset)

# AFTER
summary(student_performance_data_scale_transform)
student_performance_dataset_yield <- as.numeric(unlist(student_performance_data_scale_transform[, 99]))
hist(student_performance_dataset_yield, main = names(student_performance_data_scale_transform)[99])
```

## \<Center Data Transform\>

```{r Your Seventh Code Chunk}
#The centre data transform calculates the mean of an attribute and subtracts it from each value.

# BEFORE
summary(student_performance_dataset)
boxplot(student_performance_dataset[, 75], main = names(student_performance_dataset)[75])
boxplot(student_performance_dataset[, 76], main = names(student_performance_dataset)[76])
boxplot(student_performance_dataset[, 77], main = names(student_performance_dataset)[77])
boxplot(student_performance_dataset[, 78], main = names(student_performance_dataset)[78])
boxplot(student_performance_dataset[, 79], main = names(student_performance_dataset)[79])
boxplot(student_performance_dataset[, 80], main = names(student_performance_dataset)[80])
boxplot(student_performance_dataset[, 81], main = names(student_performance_dataset)[81])
boxplot(student_performance_dataset[, 82], main = names(student_performance_dataset)[82])
boxplot(student_performance_dataset[, 83], main = names(student_performance_dataset)[83])
boxplot(student_performance_dataset[, 84], main = names(student_performance_dataset)[84])
boxplot(student_performance_dataset[, 85], main = names(student_performance_dataset)[85])
boxplot(student_performance_dataset[, 86], main = names(student_performance_dataset)[86])
boxplot(student_performance_dataset[, 87], main = names(student_performance_dataset)[87])

model_of_the_transform <- preProcess(student_performance_dataset, method = c("center"))
print(model_of_the_transform)
student_performance_dataset_center_transform <- predict(model_of_the_transform, 
                                                        student_performance_dataset)

# AFTER
summary(student_performance_dataset_center_transform)
boxplot(student_performance_dataset_center_transform[, 75],
        main = names(student_performance_dataset_center_transform)[75])
boxplot(student_performance_dataset_center_transform[, 76],
        main = names(student_performance_dataset_center_transform)[76])
boxplot(student_performance_dataset_center_transform[, 77],
        main = names(student_performance_dataset_center_transform)[77])
boxplot(student_performance_dataset_center_transform[, 78],
        main = names(student_performance_dataset_center_transform)[78])
boxplot(student_performance_dataset_center_transform[, 79],
        main = names(student_performance_dataset_center_transform)[79])
boxplot(student_performance_dataset_center_transform[, 80],
        main = names(student_performance_dataset_center_transform)[80])
boxplot(student_performance_dataset_center_transform[, 81],
        main = names(student_performance_dataset_center_transform)[81])
boxplot(student_performance_dataset_center_transform[, 82],
        main = names(student_performance_dataset_center_transform)[82])
boxplot(student_performance_dataset_center_transform[, 83],
        main = names(student_performance_dataset_center_transform)[83])
boxplot(student_performance_dataset_center_transform[, 84],
        main = names(student_performance_dataset_center_transform)[84])
boxplot(student_performance_dataset_center_transform[, 85],
        main = names(student_performance_dataset_center_transform)[85])
boxplot(student_performance_dataset_center_transform[, 86],
        main = names(student_performance_dataset_center_transform)[86])
boxplot(student_performance_dataset_center_transform[, 87],
        main = names(student_performance_dataset_center_transform)[87])

### The Centre Basic Transform on the Crop Dataset ----
summary(student_performance_dataset)
model_of_the_transform <- preProcess(student_performance_dataset, method = c("center"))
print(model_of_the_transform)
student_performance_dataset_center_transform <- predict(model_of_the_transform, student_performance_dataset)
summary(student_performance_dataset_center_transform)
```

## \<Standardize Data Transform\>
The standardize data transform ensures that each numeric attribute has a mean value of 0 and a standard deviation of 1. This is done by combining the scale data transform and the centre data transform.
```{r Your Eighth Code Chunk}
# BEFORE
summary(student_performance_dataset)
sapply(student_performance_dataset[, 96:99], sd)

model_of_the_transform <- preProcess(student_performance_dataset,
                                     method = c("scale", "center"))
print(model_of_the_transform)
student_performance_dataset_standardize_transform <- predict(model_of_the_transform, 
                                                             student_performance_dataset)

# AFTER
summary(student_performance_dataset_standardize_transform)
sapply(student_performance_dataset_standardize_transform[, 96:99], sd)


```

## \<Normalize Data Transform\>
Normalizing a dataset implies ensuring the numerical data are between [0, 1] (inclusive).

```{r Your Ninth Code Chunk}
### The Normalize Transform on the Boston Housing Dataset ----
summary(student_performance_dataset)
model_of_the_transform <- preProcess(student_performance_dataset, method = c("range"))
print(model_of_the_transform)
student_performance_dataset_normalize_transform <- predict(model_of_the_transform, 
                                                           student_performance_dataset)
summary(student_performance_dataset_normalize_transform)


```

## \<Box-Cox Power Transform\>
The skewness informs you of the asymmetry of the distribution of results.
Similar to kurtosis, there are several ways of computing the skewness. Using
“type = 2” (discussed in a previous Lab) can be interpreted as:
 1.	Skewness between -0.4 and 0.4 (inclusive) implies that there is no
      skew in the distribution of results; the distribution of results is
      symmetrical; it is a normal distribution.
 2. Skewness above 0.4 implies a positive skew; a right-skewed distribution.
 3.	Skewness below -0.4 implies a negative skew; a left-skewed distribution.

```{r Your Tenth Code Chunk}
# BEFORE
summary(student_performance_dataset_normalize_transform)

#Calculate the skewness before the Box-Cox transform
sapply(student_performance_dataset_normalize_transform[, 96:99],  skewness, type = 2)

#Plot a histogram to view the skewness before the Box-Cox transform
hist(student_performance_dataset[, 75], main = names(student_performance_dataset)[75])
hist(student_performance_dataset[, 76], main = names(student_performance_dataset)[76])
hist(student_performance_dataset[, 77], main = names(student_performance_dataset)[77])
hist(student_performance_dataset[, 78], main = names(student_performance_dataset)[78])
hist(student_performance_dataset[, 79], main = names(student_performance_dataset)[79])
hist(student_performance_dataset[, 80], main = names(student_performance_dataset)[80])
hist(student_performance_dataset[, 81], main = names(student_performance_dataset)[81])
hist(student_performance_dataset[, 82], main = names(student_performance_dataset)[82])
hist(student_performance_dataset[, 83], main = names(student_performance_dataset)[83])
hist(student_performance_dataset[, 84], main = names(student_performance_dataset)[84])
hist(student_performance_dataset[, 85], main = names(student_performance_dataset)[85])
hist(student_performance_dataset[, 86], main = names(student_performance_dataset)[86])
hist(student_performance_dataset[, 87], main = names(student_performance_dataset)[87])

model_of_the_transform <- preProcess(student_performance_dataset, method = c("BoxCox"))
print(model_of_the_transform)
student_performance_dataset_box_cox_transform <- predict(model_of_the_transform, 
                                                         student_performance_dataset)

# AFTER
summary(student_performance_dataset_box_cox_transform)

# Calculate the skewness after the Box-Cox transform
sapply(student_performance_dataset_box_cox_transform[, 96:99],  skewness, type = 2)

#Plot a histogram to view the skewness after the Box-Cox transform
hist(student_performance_dataset_box_cox_transform[, 75],
     main = names(student_performance_dataset_box_cox_transform)[75])
hist(student_performance_dataset_box_cox_transform[, 76],
     main = names(student_performance_dataset_box_cox_transform)[76])
hist(student_performance_dataset_box_cox_transform[, 77],
     main = names(student_performance_dataset_box_cox_transform)[77])
hist(student_performance_dataset_box_cox_transform[, 78],
     main = names(student_performance_dataset_box_cox_transform)[78])
hist(student_performance_dataset_box_cox_transform[, 79],
     main = names(student_performance_dataset_box_cox_transform)[79])
hist(student_performance_dataset_box_cox_transform[, 80],
     main = names(student_performance_dataset_box_cox_transform)[80])
hist(student_performance_dataset_box_cox_transform[, 81],
     main = names(student_performance_dataset_box_cox_transform)[81])
hist(student_performance_dataset_box_cox_transform[, 82],
     main = names(student_performance_dataset_box_cox_transform)[82])
hist(student_performance_dataset_box_cox_transform[, 83],
     main = names(student_performance_dataset_box_cox_transform)[83])
hist(student_performance_dataset_box_cox_transform[, 84],
     main = names(student_performance_dataset_box_cox_transform)[84])
hist(student_performance_dataset_box_cox_transform[, 85],
     main = names(student_performance_dataset_box_cox_transform)[85])
hist(student_performance_dataset_box_cox_transform[, 86],
     main = names(student_performance_dataset_box_cox_transform)[86])
hist(student_performance_dataset_box_cox_transform[, 87],
     main = names(student_performance_dataset_box_cox_transform)[87])

```
## \<Yeo-Johnson Power Transform\>
Similar to the Box-Cox transform, the Yeo-Johnson transform reduces the skewness by shifting the distribution of an attribute and making the attribute have a more Gaussian-like distribution. The difference is that the Yeo-Johnson transform can handle zero and negative values, unlike the Box-Cox transform.

```{r Your Eleventh Code Chunk}
# BEFORE
summary(student_performance_dataset)

# Calculate the skewness before the Yeo-Johnson transform
sapply(student_performance_dataset[, 96:99],  skewness, type = 2)

# Plot a histogram to view the skewness before the Box-Cox transform
hist(student_performance_dataset[, 75], main = names(student_performance_dataset)[75])
hist(student_performance_dataset[, 76], main = names(student_performance_dataset)[76])
hist(student_performance_dataset[, 77], main = names(student_performance_dataset)[77])
hist(student_performance_dataset[, 78], main = names(student_performance_dataset)[78])
hist(student_performance_dataset[, 79], main = names(student_performance_dataset)[79])
hist(student_performance_dataset[, 80], main = names(student_performance_dataset)[80])
hist(student_performance_dataset[, 81], main = names(student_performance_dataset)[81])
hist(student_performance_dataset[, 82], main = names(student_performance_dataset)[82])
hist(student_performance_dataset[, 83], main = names(student_performance_dataset)[83])
hist(student_performance_dataset[, 84], main = names(student_performance_dataset)[84])
hist(student_performance_dataset[, 85], main = names(student_performance_dataset)[85])
hist(student_performance_dataset[, 86], main = names(student_performance_dataset)[86])
hist(student_performance_dataset[, 87], main = names(student_performance_dataset)[87])

model_of_the_transform <- preProcess(student_performance_dataset, method = c("YeoJohnson"))
print(model_of_the_transform)
student_performance_dataset_yeo_johnson_transform <- predict(model_of_the_transform, 
                                                             student_performance_dataset)

# AFTER
summary(student_performance_dataset_yeo_johnson_transform)

# Calculate the skewness after the Yeo-Johnson transform
sapply(student_performance_dataset_yeo_johnson_transform[, 96:99],  skewness, type = 2)

# Plot a histogram to view the skewness after the Box-Cox transform
hist(student_performance_dataset_box_cox_transform[, 75],
     main = names(student_performance_dataset_box_cox_transform)[75])
hist(student_performance_dataset_box_cox_transform[, 76],
     main = names(student_performance_dataset_box_cox_transform)[76])
hist(student_performance_dataset_box_cox_transform[, 77],
     main = names(student_performance_dataset_box_cox_transform)[77])
hist(student_performance_dataset_box_cox_transform[, 78],
     main = names(student_performance_dataset_box_cox_transform)[78])
hist(student_performance_dataset_box_cox_transform[, 79],
     main = names(student_performance_dataset_box_cox_transform)[79])
hist(student_performance_dataset_box_cox_transform[, 80],
     main = names(student_performance_dataset_box_cox_transform)[80])
hist(student_performance_dataset_box_cox_transform[, 81],
     main = names(student_performance_dataset_box_cox_transform)[81])
hist(student_performance_dataset_box_cox_transform[, 82],
     main = names(student_performance_dataset_box_cox_transform)[82])
hist(student_performance_dataset_box_cox_transform[, 83],
     main = names(student_performance_dataset_box_cox_transform)[83])
hist(student_performance_dataset_box_cox_transform[, 84],
     main = names(student_performance_dataset_box_cox_transform)[84])
hist(student_performance_dataset_box_cox_transform[, 85],
     main = names(student_performance_dataset_box_cox_transform)[85])
hist(student_performance_dataset_box_cox_transform[, 86],
     main = names(student_performance_dataset_box_cox_transform)[86])
hist(student_performance_dataset_box_cox_transform[, 87],
     main = names(student_performance_dataset_box_cox_transform)[87])

```

## \<Principal Component Analysis (PCA) Linear Algebra Transform\>
PCA and ICA are primarily dimensionality reduction techniques used to transform high-dimensional data into a lower-dimensional space while retaining as much variance as possible. However, they can indirectly assist in feature selection by identifying the most important features or components.

```{r Your Twelfth Code Chunk}
summary(student_performance_dataset)

model_of_the_transform <- preProcess(student_performance_dataset, method =
                                       c("scale", "center", "pca"))

print(model_of_the_transform)
student_performance_dataset_pca_dr <- predict(model_of_the_transform, student_performance_dataset)

summary(student_performance_dataset_pca_dr)


```

## \<PCA Linear Algebra Transform for Feature Extraction\>
We use the `princomp()` function is used to perform PCA on a correlation matrix.

```{r Your Thirteenth Code Chunk}
str(student_performance_dataset)
summary(student_performance_dataset)

student_performance_dataset_pca_fe <- princomp(cor(student_performance_dataset[, 96:97]))
summary(student_performance_dataset_pca_fe)

#### Scree Plot ----
# The Scree Plot shows that the 1st 2 principal components can cumulatively
# explain 92.8% of the variance, i.e., 87.7% + 5.1% = 92.8%.
factoextra::fviz_eig(student_performance_dataset_pca_fe, addlabels = TRUE)

#### Loading Values ----
# Remember: Principal components are new features created in the process of
#           dimensionality reduction. We would like to know the extent to which
#           each feature is represented in the 1st 2 principal components. We
#           can use "loading values" to determine the extent of representation.

# The loading values for each variable in the 1st 2 principal components are
# shown below:
student_performance_dataset_pca_fe$loadings[, 1:2]

# This is easier to understand using a visualization that shows the extent to
# which each variable is represented in a given component.

# In this case, it shows the extent to which each variable is represented in
# the first 2 components.


# Points to note when interpreting the visualization:
# The Cos2 value is the square cosine. It corresponds to the quality of
# representation.
#    (i) A low value means that the variable is not perfectly represented by
#         that component.
#    (ii) A high value, on the other hand, means a good representation of the
#         variable on that component.

factoextra::fviz_cos2(student_performance_dataset_pca_fe, choice = "var", axes = 1:2)

# The 8 most represented variables in the first 2 components (which we said
# represent 92.8% of the variation) are, in descending order: indus, nox,
# lstat, dis, tax, age, medv, rad

#### Biplot and Cos2 Combined Plot ----
# This can be confirmed using the following visualization.

# Points to note when interpreting the visualization:
#    (i) All the variables that are grouped together are positively correlated.
#    (ii) The longer the arrow, the better represented the variable is.
#    (iii) Variables that are negatively correlated are displayed in the
#          opposite side of the origin.

factoextra::fviz_pca_var(student_performance_dataset_pca_fe, col.var = "cos2",
                         gradient.cols = c("red", "orange", "green"),
                         repel = TRUE)

```

## \<ICA Linear Algebra Transform for Dimensionality Reduction\>
Independent Component Analysis (ICA) transforms the data to return only the independent components. The n.comp argument is required to specify the desired number of independent components. This also results in a list of attributes that are uncorrelated.


```{r Your Fourteenth Code Chunk}
if (!is.element("fastICA", installed.packages()[, 1])) {
  install.packages("fastICA", dependencies = TRUE)
}
require("fastICA")

### ICA for Dimensionality Reduction on the Boston Housing Dataset ----
summary(student_performance_dataset)

model_of_the_transform <- preProcess(student_performance_dataset,
                                     method = c("scale", "center", "ica"),
                                     n.comp = 8)
print(model_of_the_transform)
student_performance_dataset_ica_dr <- predict(model_of_the_transform, student_performance_dataset)

summary(student_performance_dataset_ica_dr)

```

**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.